### 2020-11-20 MSDS6371Project.Rmd
add rmd file

---
title: "MSDS6371Project"
author: "Allen Miller & Mingyang Nick YU"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Read in Training Data and initial inspection:
```{r}
#Load required libraries
library(tidyverse)
library(ggplot2)
library(olsrr)
#Load in Dataset
#Call original data amesHouse
amesHouse = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/train.csv",header = TRUE)
#observe data
head(amesHouse)
str(amesHouse)
#look at neighborhood
amesHouse$Neighborhood
```

#### Analysis 1:
```{r}
#extract variables we're interested for Analysis 1
data1 = amesHouse %>% select(SalePrice,GrLivArea,Neighborhood)%>%
  filter(Neighborhood=="NAmes"|Neighborhood=="Edwards"|Neighborhood=="BrkSide")
str(data1)
#383 observations selected after filtering neighborhood of "NAmes", "Edwards" and "BrkSide"
#Convert Neighborhood to factors
data1$Neighborhood = as.factor(data1$Neighborhood)
#Plot and observe the relationship between SalePrice and GrLivArea
data1 %>% ggplot(aes(x=GrLivArea,y=SalePrice))+
  geom_point()+ggtitle("Sale Price vs. Square Foot of Living area")+
  xlab("Square Foot of Living Area")+
  ylab("Sale Price")
#there are some outliers that may not have come from the same population of interest
#build model to identify the outliers
model1 = lm(SalePrice~GrLivArea,data=data1)
summary(model1)
#Look at residual plots and cook's distance
plot(model1)
cooksDistance = data.frame(distance = cooks.distance(model1))
cooksDistance%>%filter(distance>=1)
ols_plot_cooksd_bar(model1)
ols_plot_resid_stand(model1)
#ols_plot_resid_lev(model1)
# Observation 339 has cook's distance larger than 5.6, Observation 131's cook's D is larger than 1 so it may due to the face it is a unique case.
# Observation 169 and 190 has standarlized residual greater than 4, these two observations can also potentially not coming from the same population of interest
# Since our sample is sufficiently large, deleting these four outliers won't make a huge difference.
# Call these new dataset without outliers data2
data2 = data1[-c(131,169,190,339),]
str(data2)
#build new model to double check
model2 = lm(SalePrice~GrLivArea,data=data2)
summary(model2)
plot(model2)
ols_plot_cooksd_bar(model2)
ols_plot_resid_stand(model2)
#Cook's Distance within roughly 0.1, residual within 3, decent amount of observations beyond 2 (within 5%)
#Assumptions met, move on to include categorical variables
model3 = lm(SalePrice~GrLivArea+Neighborhood+GrLivArea*Neighborhood,data=data2)
summary(model3)
plot(model3)
ols_plot_cooksd_bar(model3)
ols_plot_resid_stand(model3)
#Assumptions are met, model 3 has higher Adjusted R-squared compared to model2

```